---
title: "WUP_04"
author: "Deniss Gonzalez Ramirez"
date: "20 September 2021"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)

```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  

$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
mydata<-read.csv(file="GSScsv.csv")
```

```{r}
boxplot(WORDSUM~DEGREE,data=mydata)
``` 



The middle box and the box that is located on the far right of the box plot can go completely under the box to the left labeled "graduate". The means from the two are entirely different from eachother. If you were to run a test with the two boxes ( the one in the middle and the one on the far right) it is likely that the scores on the WORDSUM test are lower than the boxes labeled "Bachelor", "Graduate", and "Junior College".


### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=mydata)
```


The individuals with degrees from a junior college, a bachelor's degree, and a graduate school degree scored a higher mean result on the WORDSUM test than those with only a high school and the junior high school degree. The numerical results above show that individuals with bachelor, junior college, or graduate school degrees have a mean of 9 or higher. The junior high school and the high school students scored a mean of 8 or less.


## Inferential Results


<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->
```{r}
model1 <- aov(WORDSUM~DEGREE,data=mydata)
summary.aov(model1)
```

The p-value is very small which signifies that the probability that the null hypothesis is true is highly unlikely; therefore we must reject the null hypothesis. There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE.


```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 15, 2, 2))
plot(TukeyHSD(model1,las=2,conf.level=.99))
```


The Tukey test above shows that HIGHSCHOOL and GRADUATE are different from each other and LT HIGHSCHOOL and GRADUATE are different from each other because 0 is not in the confidence interval and they do not cross the dotted line. 


# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->
